---
title: "Auto Standardize Long Data"
format: html
---




## ✅   The Data Pipeline 


Raw Data → Cleaning Script → Standardized Format: e.g., date, value, campus, metric, ...

Dashboard Upload → Smart Filtering → Ready for Charts

The date, value, categories structure make all chart code more intuitive:

Run Charts: aes(x = date, y = value)
Line Charts: aes(x = date, y = value, color = campus)
Control Charts: Same structure + control limits
Cohort Analysis: group_by() any categorical column



## Source automated functions

```{r}

# making automated functions available for automated data cleaning
source("automated_data_cleaning_functions.r")

```

##  Automated Data Exploration Strategy:

Load & full-luxury exploration all in one step (this is same as explore_data function)

```{r}

auto_raw_data1 <- load_and_explore("data/20250227 SCUC Snapshots 1995 to 2023-LONG.csv")

```

This gives you everything:

- Dimensions
- Column names with numbers (for easy reference)
- Data types (glimpse() output)
- First 6 rows of actual data
- Auto-detected date/numeric/categorical columns




# Full-luxury explore (if data already loaded)

```{r}

auto_raw_data2 <- explore_data(auto_raw_data1)

```




Quick exploration function (tibble of 1st 6 rows only)

```{r}

auto_raw_data3 <- quick_look(auto_raw_data1)

```


Standardize the data frame automatically (requires path to original data)
This does all that the manual pipeline does in one function

```{r}
std_data1 <- clean_eda_data("data/20250227 SCUC Snapshots 1995 to 2023-LONG.csv")

```


ALL rates and percentages must be expressed in decimal form.

If needed, convert column of your choice to decimal form

```{r}
# Usage examples:

std_data1 <- convert_column_to_decimal(std_data1, "value")
# std_data1 <- convert_column_to_decimal(std_data1, "completion_rate")
# std_data1 <- convert_column_to_decimal(std_data1, "satisfaction_score")

```



Save the results

```{r}

save_cleaned_data(std_data1, "data/auto_standardized_20250227_SCUC_Snapshots_1995_to_2023-LONG.csv")

```


